Course project submission
========================================================


```{r setup, include=FALSE}
opts_chunk$set(cache=TRUE)
``` 

Getting data into Rstudio. details are commented out
```{r get data}
# get data
total <-read.csv("pml-training.csv", na.strings=c("NA",""))
test <-read.csv("pml-testing.csv")
library(caret)
# discard NAs
totalclean <- total[,complete.cases(t(total))]
# remove non relevant variables
totalclean <- totalclean[,-c(1:7)]
# check classe frequencies in complete dataset
summary(totalclean$classe)/length(totalclean$classe)
```

Partition data
```{r partition}
set.seed(123)
inTrain <- createDataPartition(y=totalclean$classe,p=0.75, list=FALSE)
train <- totalclean[inTrain,]
valid <- totalclean[-inTrain,]
```

check that classes are ~ equally represented in the training and validation set

```{r}
summary(train$classe)/length(train$classe)
summary(valid$classe)/length(valid$classe)
```

check for presence of near-zero variance variables
```{r}
nzv <- nearZeroVar(train, saveMetrics = TRUE)
```
none of the variables are nzv
check for (and remove) higly correlated descriptors
```{r}
descrCor <- cor(train[,-53])
highlyCorDescr <- findCorrelation(descrCor, cutoff = 0.85)
trainuncorr <- train[, -highlyCorDescr]

summary(descrCor[upper.tri(descrCor)])
descrCor2 <- cor(trainuncorr[,-45])
summary(descrCor2[upper.tri(descrCor2)])
```

use doMC for parallel processing
```{r}
library(doMC)
registerDoMC(cores = 3)
```

create train control object and specify 10-fold cross validation
```{r}
ctrl1 <- trainControl(method = "cv",
                      number=10,
                      verbose = FALSE,
                      allowParallel = TRUE)
```

fit random forest to the training dataset
``` {r rftrain}
set.seed(10)
system.time({ rfmodel <- train(classe ~ ., method="rf" ,preProcess=c("center","scale"), trainControl = ctrl1, data=trainuncorr)
})
```

fit algorithm on validation set and check the out-of-sample error
``` {r validpred}
validuncorr <- valid[, -highlyCorDescr]
preds <- predict(rfmodel, validuncorr)
confusionMatrix(preds,validuncorr$classe)
```


get predictions for supplied test set
first remove same variables as the ones removed in the training (and validation) set
```{r testprep}
test <- test[,complete.cases(t(test))]
test <- test[,-c(1:7)]
test <- test[, -highlyCorDescr]
names(trainuncorr)
names(test)
```
check that the only column that differs is the last (outcome)
```{r}
setdiff(names(trainuncorr),names(test))
```
generate predictions
```{r}
predtest <- predict(rfmodel, test)
```
generate text files for submission
``` {r gentest}
setwd("texts")
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(predtest)
```

